Viscous Erosion in Porous Media

### 2018-03-21 docs/methods/erosion_methods.tex
working on edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align*}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

@@

There are several models for erosion that involve the shear
stress~\cite{par-izu2000,lag2000}.  The erosion model we adopt
prescribes that erosion occurs in the normal direction at a rate that is
proportional to the magnitude of the shear
stress~\cite{moo-ris-chi-zha-she2013}.  The constant of proportionality
only sets the time scale, so we let $\Vn = |\tau|$ as described in
equation~\eqref{eqn:erosionModel}.  For the sake of numerical stability,
we modify this law in Section~\ref{sec:thetaL}. To perform the
simulations, we freeze the geometry at each time step, solve the
incompressible Stokes equations, and then update the geometry according
to the erosion model.  The separation between these two processes is
justified since the time scale of erosion is much longer than that of
the flow.

@@

In this section, we describe the boundary integral equations employed
for the fluid equations.  In particular, we recast the incompressible
Stokes equations with Dirichlet boundary conditions 
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega,
\end{alignat*}
into an equivalent BIE.  The boundary condition $\ff$ is the prescribed
velocity $\UU$ on $\Gamma$, and it is zero on the grains $\gamma$.  In
two dimensions, a BIE for a stream function formulation is
possible~\cite{gre-kro-may1996}, but this does not extend to three
dimensions.  Therefore, we adopt a primitive variable formulation where
the velocity is written in terms of the standard double-layer
potential~\cite{lad1963,poz1992}
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\eeta$ is an unknown
density function.  As the viscosity $\mu$ also sets the time scale, we
take $\mu=1$.  \todo[inline]{Do you think we should flesh out this
time-scale stuff more precisely? i.e.~By setting $\mu = 1$ and $C = 1$
(the erosion constant), what exactly is the timescale that defines our
dimensionless time.} \todo[inline]{This might be worth doing.  Is it
something you can do quickly or do you want me to take a crack at it?}

The double-layer potential representation is only able to represent
force- and torque-free velocity fields, so we complete the velocity
representation with Stokeslets and Rotlets~\cite{pow-mir1987,pow1993}.
This is done by complementing the double-layer potential with Stokeslets
and rotlets centered inside each interior body
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ body, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.  Then,
any solution of the Stokes equations with a Dirichlet boundary condition
$\ff$ can be written in terms of the completed double-layer potential
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space resulting from the flux-free requirement
of $\ff$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the shear stress}
\label{sec:shearStressLP}
To compute the shear stress, there are two options.  First, instead of
solving the BIE~\eqref{eqn:completed_BIE} for an arbitrary density
function, a similar BIE can be solved for the
traction~\cite{mit-spa2016}.  Alternatively, the traction corresponding
to the double-layer potential of the velocity~\eqref{eqn:velocityDLP}
can be computed using appropriate layer potentials.  While the first
method bypasses a step by forming the traction jump directly, it does
not allow us to directly compute other quantities such as the vorticity,
so we use the second method.

The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss \, .
\end{align*}
The reader may notice an unusual sign convention above, but this is simply a consequence of our conventions for $\nn$ and $\ss$ (see Fig.~\ref{fig:schematic}). 
The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential, the Stokeslets, and the
rotlets.  We compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$
is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is given by
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
The deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right).
  \end{aligned}
\end{equation}
Now that the deformation tensor of each component of the completed
double-layer potential is computed, the shear stress at $\xx_0 \in
\bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

One challenge with simulating erosion is that corners develop at points
where the shear stress vanishes.  There is recent work for solving BIEs
in geometries with corners by using carefully chosen local expansions
around each corner~\cite{rac-ser2017, ser-rok2016} or panel-based graded
meshes~\cite{hel2011, gil-hao-mar2014, bre2012}.  The limitation of
these methods is that they require the corner's location, a non-uniform
spaced mesh, and some of the methods have a large condition number.  One
solution is to use two meshes---one for the quadrature the other for the
interface tracking.  Then, interpolation between the two grids is
required, and this is the approach taken by Ojala and
Tornberg~\cite{oja-tor2015}.  Instead of using two grids and
interpolation operators, we only require the uniform grid by slightly
modify the geometry to maintain sufficient regularity.  First, we add a
regularization term to the shear stress to penalize regions of high
curvature.  Second, we apply a Gaussian filter to smooth the absolute
value of the shear stress before the body shape is updated.

@@

When recast in the variables $\theta$ and $L$, the interface evolution becomes
\begin{align}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
\thalpha \Vs \right)\,.
\end{align}
These are the main evolution equations for $\theta(\alpha,t)$ and $L(t)$.
In addition, it is necessary to track a reference point for each body, so that the {\thL} variables can be converted back to cartesian coordinates. We track the coordinates of the surface mean $\mean{\xx}$, which moves according to
\begin{align}
\label{eq:xsurf}
\frac{d}{dt} \mean{\xx} = \mean{ V_s \ss + V_n \nn } \, .
\end{align}
We note that tracking the center of mass is also a valid choice, but the surface mean gives a somewhat simpler formula. Equations~\eqref{eqn:Levo1}--\eqref{eq:xsurf} comprise the main system of equations to be solved numerically in Section~\ref{sec:timeStepping}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}
There are three main numerical methods that need to be developed.  In
Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE}
is solved.  Once the density function is found,
Section~\ref{sec:shearStress} describes how the shear
stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}
and~\eqref{eqn:thetaevo1} is described. So that we can form stable
simulations with long time horizons, we use numerical methods that
achieve spectral accuracy in space, second-order accuracy in time, and
maintain an equispaced discretization of the bodies.  The boundaries are
represented in Fourier space, giving us a set of discretization points
$\xx_i$ of each body, and all derivatives and integrals (the arclength,
curvature, and mean values) are computed with spectral accuracy using
the Fourier representation.

@@

Briefly, to verity that the rescaling does not alter accuracy, we show in Table \ref{convtab} a convergence test for a 20-body simulation with the condition of fixed pressure drop imposed. The table confirms that the method retains second-order accuracy in time. In the last column, we show the computational times, indicating that the 20-body simulations require more runtime than the single-body case. Although the GMRES-iteration count is independent of the spatial resolution, $N_{\iin}$, it can depend on the number of bodies and/or their configuration. To illustrate this point, we show in Table \ref{itertab} the GMRES count for various cases ...

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  There are several models for
erosion~\cite{par-izu2000,lag2000}, but we have adopted the model that
erosion occurs in the normal direction at a rate that is proportional to
the magnitude of the shear stress~\cite{moo-ris-chi-zha-she2013}.  The
constant of proportionality only sets the time scale, so we have set
$\Vn = |\tau|$ as described in equation~\eqref{eqn:erosionModel}.  For
the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we impose the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align*}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

@@

@@

In this section, we describe an equivalent BIE formulation of the
incompressible Stokes equations with Dirichlet boundary conditions
\begin{equation}
  \begin{split}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
  \end{split}
  \label{eqn:stokes}
\end{equation}
The boundary condition $\ff$ is the prescribed velocity $\UU$ on
$\Gamma$, and it is zero on the eroding grains $\gamma$.  In two
dimensions, a BIE for a stream function formulation is
possible~\cite{gre-kro-may1996}, but this does not extend to three
dimensions.  Therefore, we adopt a primitive variable formulation where
the velocity is written in terms of the standard double-layer
potential~\cite{lad1963,poz1992}
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\eeta$ is an unknown
density function.  As the viscosity $\mu$ also sets the time scale, we
take $\mu=1$.  \todo[inline]{Do you think we should flesh out this
time-scale stuff more precisely? i.e.~By setting $\mu = 1$ and $C = 1$
(the erosion constant), what exactly is the timescale that defines our
dimensionless time.} \todo[inline]{This might be worth doing.  Is it
something you can do quickly or do you want me to take a crack at it?}

Since the double-layer potential is not able to represent force- and
torque-free velocity fields,  we complete the double-layer potential
with Stokeslets and rotlets~\cite{pow-mir1987, pow1993}
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ grainy, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.
Then, the solution of~\eqref{eqn:stokes} is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space resulting from the flux-free requirement
of $\ff$.

This BIE formulation requires that $\bd\Omega$ is smooth.
Unfortunately, smoothness is not guaranteed erosion since corners
develop at locations where the shear stress vanishes.  One option is to
modify our BIE to handle corners~\cite{rac-ser2017,
ser-rok2016, hel2011, gil-hao-mar2014, bre2012}, but these methods
require the corner's location, or a non-uniform spaced mesh, or are
ill-conditioned.  Instead, make a slight modification to the erosion
model in Section~\ref{sec:thetaL} which allows us to control the
smoothness of the grains.  In this manner, the BIE as described above
can be applied.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the shear stress}
\label{sec:shearStressLP}
Once the density function, Stokeslets, and rotlets are computed, any
desired quantity, including the shear stress, can be computed.  The
shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss \, .
\end{align*}
The reader may notice an unusual sign convention above, but this is
simply a consequence of our conventions for $\nn$ and $\ss$ (see
Fig.~\ref{fig:schematic}).  The velocity field of the completed
double-layer potential~\eqref{eqn:completed_DLP} includes velocities
from three terms: the double-layer potential, the Stokeslets, and the
rotlets.  We compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of these terms individually.

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$
is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right).
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in
\bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

@@

When recast in the variables $\theta$ and $L$, the interface evolution becomes
\begin{align}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\pderiv{\theta}{t} &= \frac{2 \pi}{L} \left( \pderiv{\Vn}{\alpha} +
\thalpha \Vs \right).
\end{align}
Inserting~\eqref{eqn:Vn2} for the normal velocity gives
\begin{align}
\label{eqn:Levo1}
\tderiv{L}{t} &= - 2 \pi \mean{\thalpha \Vn}, \\
\label{eqn:thetaevo1}
\pderiv{\theta}{t} &= \frac{2\pi}{L} \left(
\eps \mean{\atau} \ppd{\theta}{\alpha} + \pderiv{\atau}{\alpha} +
\thalpha \Vs \right)\,.
\end{align}
These are the main evolution equations for $\theta(\alpha,t)$ and $L(t)$.
In addition, it is necessary to track a reference point for each body, so that the {\thL} variables can be converted back to Cartesian coordinates. We track the coordinates of the surface mean $\mean{\xx}$, which moves according to
\begin{align}
\label{eq:xsurf}
\frac{d}{dt} \mean{\xx} = \mean{ V_s \ss + V_n \nn } \, .
\end{align}
We note that tracking the center of mass is also a valid choice, but the surface mean gives a somewhat simpler formula. Equations~\eqref{eqn:Levo1}--\eqref{eq:xsurf} comprise the main system of equations to be solved numerically in Section~\ref{sec:timeStepping}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}
\todo[inline]{Here on Bryan's sections of editing}
To perform the simulations, we freeze the geometry at each time step,
solve the incompressible Stokes equations, and then update the geometry
according to the erosion model.  The separation between these two
processes is justified since the time scale of erosion is much longer
than that of the flow.

Given this separation of time scales, there are three main numerical
methods that need to be developed.  In Section~\ref{sec:BIE}, we
describe how the BIE~\eqref{eqn:completed_BIE} is solved.  Once the
density function is found, Section~\ref{sec:shearStress} describes how
the shear stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}
and~\eqref{eqn:thetaevo1} is described. So that we can form stable
simulations with long time horizons, we use numerical methods that
achieve spectral accuracy in space, second-order accuracy in time, and
maintain an equispaced discretization of the bodies.  The boundaries are
represented in Fourier space, giving us a set of discretization points
$\xx_i$ of each body, and all derivatives and integrals (the arclength,
curvature, and mean values) are computed with spectral accuracy using
the Fourier representation.

@@

Briefly, to verity that the rescaling does not alter accuracy, we show in Table \ref{convtab} a convergence test for a 20-body simulation with the condition of fixed pressure drop imposed. The table confirms that the method retains second-order accuracy in time. In the last column, we show the computational times, indicating that the 20-body simulations require more runtime than the single-body case. Although the GMRES-iteration count is independent of the spatial resolution, $N_{\iin}$, it can depend on the number of bodies and/or their configuration. To illustrate this point, we show in Table~\ref{itertab} the GMRES count for various cases ...
\todo[inline]{preconditioning papers are~\cite{qua-cou-dar2018}
and~\cite{qua-bir2015a}}
